---
title: "Analysis: A common neural mechanism underlies experiences of passage of time"
author:
- Mathis Jording, Mark Wünsche, Megha Madathil, David H.V. Vogel,
- Shivakumar Viswanathan, Kai Vogeley
date: "`r format(Sys.time(), '%d.%m.%y')`"
output:
  pdf_document:
    keep_tex: true
  pdf: default
header-includes:
- \usepackage{float}
- \usepackage{fvextra}
- \usepackage{color}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{ commandchars=\\\{\}, breaklines,
  breakanywhere, fontsize=\small }
- \DefineVerbatimEnvironment{verbatim}{Verbatim}{ breaklines, breakanywhere, fontsize=\small
  }
params:
  output_dir: D:/EEGST data/output
---


<!--- Markdown settings -->
```{r setup, include=FALSE}

# set size of figures
fig_width = 160
fig_height = 100

knitr::opts_chunk$set(echo = FALSE,
                      fig.align = 'center',
                      fig_crop = FALSE,
                      fig.pos = "H",
                      fig.width = fig_width / 25.4,
                      fig.height = fig_height / 25.4)

```

<!--- Packages -->
```{r Packages, include = FALSE, message=FALSE, warning=FALSE}
# Change 'message=TRUE' & 'warning=TRUE' for messages and warnings

# Core
library("knitr")
library("grid")

# Modeling / statistics
library("lme4")
library("emmeans")
library("psych")

# Data manipulation
library("plyr")
library("tidyverse")  # loads dplyr, ggplot2, tibble, etc.

# Plotting
library("ggsignif")
library("gridExtra")
library("ggh4x")
library("scales")
library("ggforce")
library("gridExtra")
library("egg")
library("gtable")

# Utilities
library("Rmisc")
```

<!--- Directories and settings -->
```{r directories, results="hide"}
# data folders
#dir.data = file.path("D:", "EEGST data")

dir.data = file.path("C:", "Eigene Dateien", "1. Projekte", "Virtualtimes",
                     "Projekte und Studien", "sfct VR EEG", "Manuscript & publication",
                     "white paper", "upload", "data")

# output folders (this can be changed if output should be stored somewhere else)
#dir.output = file.path(dir.data, "output")

dir.output = file.path("C:", "Eigene Dateien", "1. Projekte", "Virtualtimes",
                       "Projekte und Studien", "sfct VR EEG", "Manuscript & publication",
                       "white paper", "upload", "output")

```

<!--- Settings for plot aesthetics -->
```{r output_settings, include=FALSE}

# Plot settings
print_as = ".png"
dpi_pngs = 300
height_prediction_plots = 8

# Color for RS - Starfield comparison
colors_rs = "#d6b900"
colors_sf_diff = "#00c1ffff"
colors_sf_same = "#bc11a0ff"

# Colors for non-matching sf pairs
colors_diff = "#888888"
# Colors for matching levels of velocity / density
colors_vel = c("#ff9666", "#ff5100", "#993000")
colors_den = c("#28cc00", "#146600")

# Colors for Validation - generalization comparisons
colors.val_gen = c("dodgerblue", "darkorange")


# position dodge for PoTJ plots
dodge_potj = .3

# Position dodge for validation / crossvalidation plots
dodge_val_gen = .25




# Aesthetics
theme_grid = theme(plot.title = element_text(size=10, face ="bold", family = "sans",
                                             margin = margin(t = 0, r = 0, b = 20, l = 0),
                                             hjust = 0.5),
                   axis.title.x = element_text(size = 9),
                   axis.title.y = element_text(size = 9),
                   axis.text.x = element_text(size = 8),
                   axis.text.y = element_text(size = 8),
                   
                   axis.ticks = element_line(linewidth = 0.05),
                   axis.line = element_line(colour = "black"),
                   plot.margin = margin(t = 2, r = 2, b = 2, l = 2,
                                        unit = "mm"),
                   # legend
                   legend.title = element_text(size = 9),
                   legend.text = element_text(size = 8),
                   legend.key.size = unit(4, "mm"),
                   
                   #legend.margin = margin(0, 0.01, 0, 0, unit = "npc"),
                   #legend.box.margin = margin(0,0.005,0,0, unit= "npc"),
                   # panel and background
                   panel.background = element_blank(),
                   panel.grid.major.x = element_blank(),
                   panel.grid.minor.x = element_blank(),
                   panel.grid.major.y = element_line(linewidth=.05, color="lightgrey"), 
                   panel.spacing = unit(2, "lines"))


# setting number of digits for all prints to 2
options(digits = 3)
```

<!--- Functions -->
```{r function_p_annotation(), include = FALSE}

# p_annotation(): Function that outputs annotations of significance level as ***, **, * or 'n.s.'

p_annotation <- function(p_value){
  if (p_value <.001) {
    p_annot = "***"
  } else if (p_value <.01) {
    p_annot = "**"
  } else if (p_value <.05) {
    p_annot = "*"
  } else {
    p_annot = "n.s."
  }
  return(p_annot)
}
```

<!--- Loading data -->
```{r loading_data, include = FALSE, results="hide"}

# load dataframe 'data.subjects'
load(file = file.path(dir.data, "data.subjects.Rda"))

# load dataframe 'data.subjects_excluded'
load(file = file.path(dir.data, "data.subjects_excluded.Rda"))

# load dataframe 'data.trials'
load(file = file.path(dir.data, "data.trials.Rda"))




```



# Descriptive Statistics

## Data availability
```{r data_availability, comment = ""}

excluded_missing_data = subset(data.subjects_excluded, available_nc_snapshots_pro <0.75)
excluded_subjects = unique(data.subjects_excluded$subject)

# excluded_missing_data = subset(data.subjects_allcon, available_nc_snapshots_pro <0.75)
# excluded_subjects = unique(excluded_missing_data$subject)

# Build statement about subjects excluded due to missing data
sample_exclusion_statement = ""
if (length(excluded_subjects) > 0) {
  sample_exclusion_statement = paste("Excluded subjects due to missing data:", length(excluded_subjects))
  
  for (i in 1:length(excluded_subjects)){
    subject = excluded_missing_data[i,]
    
    subject_exclusion_statement = paste("  ", i, ": ",
                                        "available snapshots: ", subject$available_nc_snapshots, " / 540 ",
                                        "(", round(subject$available_nc_snapshots_pro*100, 2), "%)", sep = "")
    
    sample_exclusion_statement = paste(sample_exclusion_statement,
                                       subject_exclusion_statement,
                                       sep = "\n")
  }
}else {subject_exclusion_statement = "No subjects excluded due to missing data"}


# Available rs snapshots remaining subjects
available_rs_statement = paste("  - remaining RS snapshots: M = ", 
                               round(mean(data.subjects$available_rs_snapshots),2),
                               " / 300 (",
                               round(mean(data.subjects$available_rs_snapshots_pro * 100),2),
                               "%; SD = ",
                               round(sd(data.subjects$available_rs_snapshots),2),
                               ", min = ",
                               min(data.subjects$available_rs_snapshots),
                               ", max = ",
                               max(data.subjects$available_rs_snapshots),
                               ")\n     -> In total, ",
                               sum(data.subjects$available_rs_snapshots),
                               " RS snapshots available for analysis",
                               sep = "")

# Available rs snapshots remaining subjects
available_sf_statement = paste("  - remaining SF snapshots: M = ", 
                               round(mean(data.subjects$available_nc_snapshots),2),
                               " / 540 (",
                               round(mean(data.subjects$available_nc_snapshots_pro * 100),2),
                               "%; SD = ",
                               round(sd(data.subjects$available_nc_snapshots),2),
                               ", min = ",
                               min(data.subjects$available_nc_snapshots),
                               ", max = ",
                               max(data.subjects$available_nc_snapshots),
                               ")\n     -> In total, ",
                               sum(data.subjects$available_nc_snapshots),
                               " SF snapshots available for analysis",
                               sep = "")


# print statements about data availability
cat(paste(sample_exclusion_statement,
          "\n\n",
          "Remaining subjects: ", sum(data.subjects$included == TRUE), "\n",
          available_rs_statement,
          "\n",
          available_sf_statement,
          sep = ""))


```

## Demografic data
```{r descriptive_statistics, comment = ""}

# building dataframe for demografic data with one row per subject
sample_info = data.subjects[,c('subject', 'gender', 'age', 'handedness', 'aq', 'bdi', 'css')]

# Compute descriptive statistics of group info as dataframe
group_stats = describe(sample_info)

# List occurrences of gender
gender_table = table(sample_info$gender)
# List occurrences of handedness
handedness_table = table(sample_info$handedness)

# print gender, handedness and age info for group
cat(paste(gender_table[1], " Participants identified as female, ", gender_table[2], " as male\n",
          handedness_table["right"], " Participants identified as right-handed, ", handedness_table["left"], " as left-handed\n",
          "Age: mean = ", round(group_stats["age", "mean"],2), ", sd = ", round(group_stats["age", "sd"],2), "; ",
          "min = ", group_stats["age", "min"], ", max = ", group_stats["age", "max"], sep = ""))

```



\newpage
# Behavioral analysis

## Checking color rating accuracies
```{r color_rating_acc, comment = ""}
# for each trial keep the last entry (substring) from the color log to determine the last color (change)
last_colorchange_per_trial <- unlist(lapply(strsplit(data.trials$color_log, ","), tail, n = 1L))

# for each color (change) keep the last letter to determine the predominant color (r, g, b)
last_color_per_trial <- unlist(lapply(strsplit(last_colorchange_per_trial, ":"), tail, n = 1L))

# for each trial, keep only the first letter from the answer to the color rating (r, g, b)
colorrating <- unlist(lapply(substr(data.trials$color_rating, start = 1, stop = 1), tolower))

# for each trial, compare predominant color and color rating 
data.trials$colorrating_correct = last_color_per_trial == colorrating

# for each subject compute the over all color rating accuracy
data.trials.ratingaccuracy = data.trials %>%
  dplyr::group_by(subject) %>%
  dplyr::summarize(accuracy = mean(colorrating_correct))

# add color rating accuracy information to data.subjects
data.subjects <- merge(x = data.subjects, y = data.trials.ratingaccuracy, by = "subject", all.x = TRUE)

# Print information about color tracking accuracy
cat(paste("Color rating accuracy over whole sample: ",
          "M = ", round(describe(data.subjects$accuracy)[[3]],2), ", SD = ", round(describe(data.subjects$accuracy)[[4]],2),
          "; Min = ", describe(data.subjects$accuracy)[[8]],  ", Max = ", describe(data.subjects$accuracy)[[9]],
          sep = ""))
```


## Effect of velocity and density on potjs
```{r analysing_PoTJs, fig.cap = "Participant-wise PoTJs (y-axis) averaged for each variant of Velocity (x-axis) and Density (colors). Participants averages are depicted as small, transparent circles, group averages per level are depicted as bigger opaque circles, 95% within-participant confidence intervals are shown as error bars. Pairwise comparisons revealed significant differences between all variants with different Velocity levels and between low and high Density when Velocity was slow or fast but not zero (* < 0.05; ** < 0.01, *** < 0.001, n.s.: not significant)."}
# Likelihood-Ratio test for effects of velocity and density 
lms.0 <- lmer(PoTJ ~ 1 + (1 | subject), data = data.trials, REML = FALSE)
lms.vel <- lmer(PoTJ ~ as.numeric(velocity) + (1 | subject), data = data.trials, REML = FALSE)
lms.velden <- lmer(PoTJ ~ as.numeric(velocity) + as.numeric(density) + (1 | subject), data = data.trials, REML = FALSE)
lms.velxden <- lmer(PoTJ ~ as.numeric(velocity) * as.numeric(density) + (1 | subject), data = data.trials, REML = FALSE)
anova.potj = anova(lms.0, lms.vel, lms.velden, lms.velxden)

# Conduct pairwise comparison between all condition combinations (tukey-test, 'emmeans'-package) --> suppl
lms.velxden <- lmer(PoTJ ~ velocity * density + (1 | subject), data = data.trials)
pairwise_results_all = lsmeans(lms.velxden, pairwise ~ velocity:density, pbkrtest.limit = 3150)

# Save tables as dataframes
table_anova <- as.data.frame(anova.potj)
table_coefficients <- as.data.frame(summary(lms.velxden)$coefficients)
table_pairwise <- as.data.frame(pairwise_results_all$contrasts)

saveRDS(
  list(anova = table_anova,
       coefficients = table_coefficients,
       pairwise = table_pairwise
  ),
  file = file.path(dir.output, "tables", "tables_potj.rds")
)

# Printing results as tables
knitr::kable(table_anova,
             caption = "Likelihood-Ratio test for effects of velocity and density")

knitr::kable(table_coefficients,
             caption = "Model summary velxden")

knitr::kable(table_pairwise,
             caption = "Pairwise comparisons of PoTJs (Tukey corrected)",
             digits = 3)
```

<!--- Plotting PoTJs -->
```{r plotting_PoTJs, fig.cap = "Participant-wise PoTJs (y-axis) averaged for each variant of Velocity (x-axis) and Density (colors). Participants averages are depicted as small, transparent circles, group averages per level are depicted as bigger opaque circles, 95% within-participant confidence intervals are shown as error bars. Pairwise comparisons revealed significant differences between all variants with different Velocity levels and between low and high Density when Velocity was slow or fast but not zero (* < 0.05; ** < 0.01, *** < 0.001, n.s.: not significant)."}
# create dataframe with participant means (per condition combination)
data.trials_avg <- data.trials %>%
  dplyr::group_by(subject, velocity, density) %>%
  dplyr::summarize(mean_potj = mean(PoTJ), .groups = "drop")

# compute within-subject cls
dfwc_potj_vel <- summarySEwithin(data.trials, measurevar="PoTJ", withinvars=c("velocity", "density"),
                        idvar="subject", na.rm=TRUE, conf.interval=.95)

# Effects of velocity on PoTJs in controls
plot_potjs = ggplot(data=data.trials, aes(x = velocity, y = PoTJ, group = density, color = density)) +
  geom_point(data = data.trials_avg, aes (x = velocity, y = mean_potj, group = density, color = density, fill = density),
             shape=16, size = 2, alpha = .2, position = position_dodge(width = dodge_potj)) +
  xlab("Velocity")+
  ylab("PoTJ")+
  geom_errorbar(data = dfwc_potj_vel, width=.3, aes(ymin=PoTJ-ci, ymax=PoTJ+ci), position = position_dodge(width = dodge_potj)) +
  geom_point(data = dfwc_potj_vel, shape=21, size=3, aes(x = velocity, y = PoTJ, group = density, fill = density), position = position_dodge(width = dodge_potj)) +
  theme_grid+
  scale_y_continuous(limits=c(0,1.1), breaks=seq(0,1, by = .25))+
  coord_cartesian(expand=FALSE, clip="off")

# add annotations for pairwise comparisons
plot_potjs <- plot_potjs +
  # add lines
  annotate("segment", x = c(1, 2, (1-(dodge_potj/4)), (2-(dodge_potj/4)), (3-(dodge_potj/4))), xend = c(2, 3, (1+(dodge_potj/4)), (2+(dodge_potj/4)), (3+(dodge_potj/4))),
           y = c(1.07, 1.07, 0.97, 0.97, 0.97), yend = c(1.07, 1.07, 0.97, 0.97, 0.97),
           colour = "black", linewidth=0.5)+
  # add asterics annotations
  annotate("text", x = c(1.5, 2.5, 2, 3), y = c(1.09, 1.09, 0.99, 0.99),
           label = c("***", "***", "*", "*"), size = 6)+

  # add 'n.s.' annotations (needs smaller font)
  annotate("text", x = c(1), y = c(1.01),
           label = c("n.s."), size = 4)+
  # add ticks
  annotate("segment", x = c(1, 2, 3, (1-(dodge_potj/4)), (1+(dodge_potj/4)), (2-(dodge_potj/4)), (2+(dodge_potj/4)), (3-(dodge_potj/4)), (3+(dodge_potj/4))),
           xend = c(1, 2, 3, (1-(dodge_potj/4)), (1+(dodge_potj/4)), (2-(dodge_potj/4)), (2+(dodge_potj/4)), (3-(dodge_potj/4)), (3+(dodge_potj/4))),
         y = c(1.07, 1.07, 1.07, 0.97, 0.97, 0.97, 0.97, 0.97, 0.97), yend = c(1.05, 1.05, 1.05, 0.95, 0.95, 0.95, 0.95, 0.95, 0.95),
         color = "black", linewidth=0.5)

ggsave(plot = plot_potjs,
       filename = "fig2_potjs.png", path = dir.output,
       width = fig_width, height = fig_height * 0.8, units = "mm", dpi = dpi_pngs)

plot_potjs
```
**$\Rightarrow$ Passage of time experience varies with starfield visual features**

\newpage
# EEG analysis

## Occipital power differences between RS and SF

```{r data_face_validity, include=TRUE}

# Load dataframe 'data.epochs.occamp'
load(file = file.path(dir.data, "data.epochs.occamp.Rda"))

# statistics
lms.postamp.0 <- lmer(posterior_meanamp ~ 1 + (1 | subject), data = data.epochs.occamp, REML = FALSE)
lms.postamp.1 <- lmer(posterior_meanamp ~ condition + (1 | subject), data = data.epochs.occamp, REML = FALSE)
anova.postamp = anova(lms.postamp.0, lms.postamp.1)

# Save tables as dataframes
table_anova <- as.data.frame(anova.postamp)
table_coefficients <- as.data.frame(summary(lms.postamp.1)$coefficient)

saveRDS(
  list(anova = table_anova,
       coefficients = table_coefficients
  ),
  file = file.path(dir.output, "tables", "tables_postamp.rds")
)

# Printing results as tables
knitr::kable(table_anova,
             caption = "Likelihood-Ratio test mean amplitude in posterior channels")

knitr::kable(table_coefficients,
             caption = "Model summary postamp.1")

```

**$\rightarrow$ Reduced alpha power in the task-state compared to RS in occiptial channels**

\newpage


## Epoch feature correlations
```{r feat_corr_rs_sf}

# load dataframe "group_corrs"
load(file = file.path(dir.data, "epoch_feature_correlations.Rda"))


# exclude diagonal (correlations between epochs of same type (same velocity & same density))
group_corrs_norspairs = subset(group_corrs, !(comparison == "RS - RS"))


# =====
# Analysis

# rs vs. starfield
lms.featurecorrelations.type.0 <- lmer(corr ~ 1 + (1 | subject), data = group_corrs_norspairs, REML = FALSE)
lms.featurecorrelations.type.1 <- lmer(corr ~ 1 + comparison + (1 | subject), data = group_corrs_norspairs, REML = FALSE)
anova.featurecorrelations.type <- anova(lms.featurecorrelations.type.0, lms.featurecorrelations.type.1)

# Conduct pairwise comparison (tukey-test, 'emmeans'-package) --> suppl
pairwise_comparison_similarity = lsmeans(lms.featurecorrelations.type.1, pairwise ~ comparison)




# Save tables as dataframes
table_anova <- as.data.frame(anova.featurecorrelations.type)
table_pairwise <- as.data.frame(pairwise_comparison_similarity$contrasts)

saveRDS(
  list(anova = table_anova,
       coefficients = table_coefficients,
       pairwise = table_pairwise
  ),
  file = file.path(dir.output, "tables", "tables_featurecorrelations.rds")
)



# Printing tables
knitr::kable(table_anova,
             caption = "Likelihood-Ratio of feature correlations")

knitr::kable(table_pairwise$contrasts,
             caption = "Pairwise comparisons of feature similarity (Tukey corrected)",
             digits = 3)
```



```{r plotting_feat_corrs_rs_sf, fig.cap = "Boxplots of epoch feature correlations showing the significantly increasing similarity from RS – SF pairs (yellow) to SF pairs of different variants (blue) to SF pairs of the same variant (purple). Circles: Participant-wise averaged epoch feature correlation. Boxplots: middle line: median; hinges: 25% - 75%-quantile; whiskers: smallest value below 25%-quantile and largest values above 75%-quantile within 1.5 x inter-quartile-range. Significance levels of t-tests: * < 0.05; ** < 0.01, *** < 0.001, n.s.: not significant."}
# =====
# Plotting


p_value.rs_sf_diff = broom::tidy(pairwise_comparison_similarity$contrasts)[1,"adj.p.value"]
p_annot.rs_sf_diff = p_annotation(p_value.rs_sf_diff)

p_value.sf_diff_sf_same = broom::tidy(pairwise_comparison_similarity$contrasts)[3,"adj.p.value"]
p_annot.sf_diff_sf_same = p_annotation(p_value.sf_diff_sf_same)

ylimits = c(0.6, 0.95)

# prepare dataframe for plotting
plotting_data.task = group_corrs_norspairs %>%
  dplyr::group_by(subject, comparison) %>%
  dplyr::summarize(corr = mean(corr), .groups = "drop")


plot_feature_corrs.task = ggplot(data= plotting_data.task, aes(x = comparison, y = corr)) +
  geom_point(aes(color = comparison),
             shape=16, size = 3, alpha = .3) +
  geom_line(aes(group = subject), alpha = .1, color = "black")+
  geom_boxplot(width = .3, alpha = 0) +
  ylim(ylimits)+
  geom_signif(comparisons = list(c("RS - SF", "Diff. SF")),
              annotation = c(p_annot.rs_sf_diff), tip_length = 0.03,
              margin_top = 0.1, extend_line = -.005,
              color = "black", textsize = 6)+
  geom_signif(comparisons = list(c("Diff. SF", "Same SF")),
              annotation = c(p_annot.sf_diff_sf_same), tip_length = 0.03,
              margin_top = 0.1, extend_line = -.005,
              color = "black", textsize = 6)+
  scale_fill_manual(values=c(colors_rs, colors_sf_diff, colors_sf_same))+
  scale_color_manual(values=c(colors_rs, colors_sf_diff, colors_sf_same))+
  theme_grid+
  theme(legend.position="none",
        axis.title.x=element_blank())


ggsave(plot = plot_feature_corrs.task,
       filename = "fig4b_feature_corrs_task.png", path = dir.output,
       width = fig_width / 3, height = fig_height, units = "mm", dpi = dpi_pngs)

plot_feature_corrs.task
```

**$\rightarrow$ Starfield epochs more similar to other starfield epochs than to RS epochs**

**$\rightarrow$ Starfield epochs from same variant more similar than from different variants**

\newpage


```{r feat_corr_sf_variants}
# -----
# VELOCITY

# comparing between different pairs and same velocity pairs
group_corrs.same_vel = subset(group_corrs, sameden == FALSE)

lms.featurecorrelations.same_vel.0 <- lmer(corr ~ 1 + (1 | subject), data = group_corrs.same_vel, REML = FALSE)
lms.featurecorrelations.same_vel.1 <- lmer(corr ~ 1 + samevel + (1 | subject), data = group_corrs.same_vel, REML = FALSE)
anova.featurecorrelations.same_vel <- anova(lms.featurecorrelations.same_vel.0, lms.featurecorrelations.same_vel.1)

kable(anova.featurecorrelations.same_vel, caption = "Likelihood-Ratio Test: comparing epoch feature similarity between different and same velocity", digits = 3)



# comparing between zero, low and high velocity pairs
group_corrs.vel = subset(group_corrs.same_vel, samevel == TRUE)

lms.featurecorrelations.vel.0 <- lmer(corr ~ 1 + (1 | subject), data = group_corrs.vel, REML = FALSE)
lms.featurecorrelations.vel.1 <- lmer(corr ~ 1 + as.numeric(velocity) + (1 | subject), data = group_corrs.vel, REML = FALSE)
anova.featurecorrelations.vel <- anova(lms.featurecorrelations.vel.0, lms.featurecorrelations.vel.1)

kable(anova.featurecorrelations.vel, caption = "Likelihood-Ratio Test: comparing epoch feature similarity between velocity variants", digits = 3)


# -----
# DENSITY

# comparing between different pairs and same density pairs
group_corrs.same_den = subset(group_corrs, samevel == FALSE)

lms.featurecorrelations.same_den.0 <- lmer(corr ~ 1 + (1 | subject), data = group_corrs.same_den, REML = FALSE)
lms.featurecorrelations.same_den.1 <- lmer(corr ~ 1 + sameden + (1 | subject), data = group_corrs.same_den, REML = FALSE)
anova.featurecorrelations.same_den <- anova(lms.featurecorrelations.same_den.0, lms.featurecorrelations.same_den.1)

kable(anova.featurecorrelations.same_den, caption = "Likelihood-Ratio Test: comparing epoch feature similarity between different and same density", digits = 3)


# comparing between zero, low and high density pairs
group_corrs.den = subset(group_corrs.same_den, sameden == TRUE)

lms.featurecorrelations.den.0 <- lmer(corr ~ 1 + (1 | subject), data = group_corrs.den, REML = FALSE)
lms.featurecorrelations.den.1 <- lmer(corr ~ 1 + as.numeric(density) + (1 | subject), data = group_corrs.den, REML = FALSE)
anova.featurecorrelations.den <- anova(lms.featurecorrelations.den.0, lms.featurecorrelations.den.1)

kable(anova.featurecorrelations.den, caption = "Likelihood-Ratio Test: comparing epoch feature similarity between density variants", digits = 3)



# Save tables as dataframes
table_anova.same_vel <- as.data.frame(anova.featurecorrelations.same_vel)
table_anova.vel <- as.data.frame(anova.featurecorrelations.vel)
table_anova.same_den <- as.data.frame(anova.featurecorrelations.same_den)
table_anova.den <- as.data.frame(anova.featurecorrelations.den)


saveRDS(
  list(anova_same_vel = table_anova.same_vel,
       anova_vel = table_anova.vel,
       anova_same_den = table_anova.same_den,
       anova_den = table_anova.den
  ),
  file = file.path(dir.output, "tables", "tables_featcorrs_variants.rds")
)


```


```{r plotting_feat_corr_sf_variants, fig.cap = "Boxplots of epoch feature correlations showing a significant increase of similarity for same velocity SF pairs (orange) and same density SF pairs (green) compared to different velocity and density pairs (grey) as well as a significantly increased similarity for higher velocities pairs compared to lower velocity pairs (light to dark orange). Circles: Participant-wise averaged epoch feature correlation. Boxplots: middle line: median; hinges: 25% - 75%-quantile; whiskers: smallest value below 25%-quantile and largest values above 75%-quantile within 1.5 x inter-quartile-range. Significance levels of t-tests: * < 0.05; ** < 0.01, *** < 0.001, n.s.: not significant."}
# ----------
# plotting - combined plot

# delete all correlations with rs epochs
plotting_data.corrs_conditions = subset(group_corrs, !(cond_a == "rs" | cond_b == "rs"))

# delete all correlations where velocity and density are the same on both sides (diagonal)
plotting_data.corrs_conditions = subset(plotting_data.corrs_conditions, !(samevel & sameden))

# add variable for stimulation intensity
plotting_data.corrs_conditions$intensity = rep(0, nrow(plotting_data.corrs_conditions))
plotting_data.corrs_conditions$intensity[plotting_data.corrs_conditions$samevel] = plotting_data.corrs_conditions$velocity[plotting_data.corrs_conditions$samevel]
plotting_data.corrs_conditions$intensity[plotting_data.corrs_conditions$sameden] = plotting_data.corrs_conditions$density[plotting_data.corrs_conditions$sameden]

# prepare dataframe for plotting
plotting_data.corrs_conditions = plotting_data.corrs_conditions %>%
  dplyr::group_by(subject, similarity, intensity) %>%
  dplyr::summarize(corr = mean(corr), .groups = "drop")

#
plotting_data.corrs_conditions$similarity = factor(plotting_data.corrs_conditions$similarity,
                                                   labels = c("Diff. Vel. & Den.", "Same Velocity", "Same Density"))

# create new variable with unique codes for each condition combination (similarity * intensity) 
plotting_data.corrs_conditions$combinedconditions = interaction(as.ordered(plotting_data.corrs_conditions$intensity), plotting_data.corrs_conditions$similarity)
plotting_data.corrs_conditions$combinedconditions = droplevels(plotting_data.corrs_conditions$combinedconditions)


plotting_data.corrs_conditions$combinedconditions <- as.character(plotting_data.corrs_conditions$combinedconditions)

new_levels = c(plotting_data.corrs_conditions$combinedconditions[1], "GAP1", plotting_data.corrs_conditions$combinedconditions[2:4], "GAP2", plotting_data.corrs_conditions$combinedconditions[5:6])


new_labels = c("",
               " ",
               "zero", "slow", "fast",
               "  ",
               "few", "many")

plotting_data.corrs_conditions$combinedconditions <- factor(plotting_data.corrs_conditions$combinedconditions, levels = new_levels, labels = new_labels)


plot_feature_corrs_conditions <- ggplot(data = plotting_data.corrs_conditions, aes(x = combinedconditions, y = corr)) +
  geom_point(aes(color = combinedconditions),
             shape=16, size = 3, alpha = .3) +
  geom_line(aes(group = subject), alpha = .1, color = "black")+
  geom_boxplot(width = .4, alpha = .1) +
  
  # colors
  scale_fill_manual(values=c(colors_diff, colors_vel[1], colors_vel[2], colors_vel[3], colors_den[1], colors_den[2])) +
  scale_color_manual(values=c(colors_diff, colors_vel[1], colors_vel[2], colors_vel[3], colors_den[1], colors_den[2])) +
  
  scale_x_discrete(drop = FALSE, axis_labels <- c("Cond1", "", "Cond2", "Cond3", "Cond4", " ", "Cond5", "Cond6"))+
  coord_cartesian(clip = "off") +
  theme_grid +
  theme(legend.position="none",
        strip.background = element_blank(),
        strip.placement = "outside",
        axis.title.x=element_blank(),
        plot.margin = margin(t = 2, r = 2, b = 7, l = 2,
                                        unit = "mm"),
        axis.ticks.length = unit(0, "pt"))


# -----
# Add markers for statistical tests

# determine annotation for significance levels
p_annot.same_vel = p_annotation(anova.featurecorrelations.same_vel$"Pr(>Chisq)"[2])
p_annot.vel = p_annotation(anova.featurecorrelations.vel$"Pr(>Chisq)"[2])
p_annot.same_den = p_annotation(anova.featurecorrelations.same_den$"Pr(>Chisq)"[2])
p_annot.den = p_annotation(anova.featurecorrelations.den$"Pr(>Chisq)"[2])


plot_feature_corrs_conditions <- plot_feature_corrs_conditions +
  annotate("segment", x = c(1, 3, 1, 7), xend = c(4, 5, 7.5, 8), y = c(0.94, 0.92, 0.97, 0.95), yend = c(0.94, 0.92, 0.97, 0.95), colour = "black", linewidth=0.5)+
  annotate("text", x = c(2.5, 4, 4.25), y = c(0.943, 0.923, 0.973), label = c(p_annot.same_vel, p_annot.vel, p_annot.same_den), size = 5)+
  annotate("text", x = c(7.5), y = c(0.958), label = c(p_annot.den), size =3.5)
    
  
# Add 2nd line of axis labels for similarity group
plot_feature_corrs_conditions <- plot_feature_corrs_conditions+
  annotation_custom(
    grob = textGrob("Diff. Vel. & Den.", gp = gpar(fontsize = 9)),
    xmin = 1, xmax = 1, ymin = 0.65, ymax = 0.65)+
  annotation_custom(
    grob = textGrob("Same Velocity", gp = gpar(fontsize = 9)),
    xmin = 4, xmax = 4, ymin = 0.65, ymax = 0.65)+
  annotation_custom(
    grob = textGrob("Same Density", gp = gpar(fontsize = 9)),
    xmin = 7.5, xmax = 7.5, ymin = 0.65, ymax = 0.65)

plot_feature_corrs_conditions

ggsave(plot = plot_feature_corrs_conditions,
       filename = "fig4c_feature_corrs_variants.png", path = dir.output,
       width = fig_width * 0.65, height = fig_height, units = "mm", dpi = dpi_pngs)

```

**$\rightarrow$ Increasing similarity between epochs with increasing velocity but not with increasing density**

**$\Rightarrow$ Starfield variants induce distinctive oscillation states**

\newpage


## Prediction of PoT between velocity / density levels 

```{r predictions_between_indv_levels, comment = ""}

# load dataframe 'data_vel' containing validation & generalization between velocity levels 
load(file = file.path(dir.data, "data_vel.Rda"))

# create dataframe with averages over all velocity levels
data_vel_avg = data_vel %>%
  dplyr::group_by(subject, type) %>%
  dplyr::summarize(z = mean(z, na.rm = TRUE), .groups = "drop")
data_vel_avg$vel = "total"


# load dataframe 'data_den' containing validation & generalization between density levels 
load(file = file.path(dir.data, "data_den.Rda"))

# create dataframe with averages over all density levels
data_den_avg = data_den %>%
  dplyr::group_by(subject, type) %>%
  dplyr::summarize(z = mean(z, na.rm = TRUE), .groups = "drop")
data_den_avg$den = "total"


# =====
# Analysis

# -----
# Analyse results from validation / generalization between velocity levels

# compute average and perform t-test for validation / generalization
ttest_validation_vel = t.test(data_vel_avg[data_vel_avg$type == "validation",]$z)
ttest_generalization_vel = t.test(data_vel_avg[data_vel_avg$type == "generalization",]$z)

ttest_validation_den = t.test(data_den_avg[data_den_avg$type == "validation",]$z)
ttest_generalization_den = t.test(data_den_avg[data_den_avg$type == "generalization",]$z)


#subset(data_vel, type = "generalization")

# analyse differential effects of velocity levels
lm_vel0 = lmer(z ~ 1 + (1 | subject), subset(data_vel, type == "generalization"), REML = FALSE)
lm_vel1 = lmer(z ~ as.factor(vel) + (1 | subject), subset(data_vel, type == "generalization"), REML = FALSE)
anova_vel = anova(lm_vel0, lm_vel1)

# Conduct pairwise comparison between velocity levels (tukey-test, 'emmeans'-package)
pairwise_comparison_vel = lsmeans(lm_vel1, pairwise ~ vel)

# analyse differential effects of density levels
lm_den0 = lmer(z ~ 1 + (1 | subject), subset(data_den, type == "generalization"), REML = FALSE)
lm_den1 = lmer(z ~ as.factor(den) + (1 | subject), subset(data_den, type == "generalization"), REML = FALSE)
anova_den = anova(lm_den0, lm_den1)

# Conduct pairwise comparison between density (tukey-test, 'emmeans'-package) 
pairwise_comparison_den = lsmeans(lm_den1, pairwise ~ den)



# =====
# Printing

cat(paste("Testing cross-validation results within velocity conditions against zero:\n",
          "t(", ttest_validation_vel[[2]], ") = ", ttest_validation_vel[[1]], ", p = ", ttest_validation_vel[[3]],
          "\n\n",
          "Testing generalization results between velocity conditions against zero:\n",
          "t(", ttest_generalization_vel[[2]], ") = ", ttest_generalization_vel[[1]], ", p = ", ttest_generalization_vel[[3]],
          "\n\n\n",
          "Testing cross-validation results within density conditions against zero:\n",
          "t(", ttest_validation_den[[2]], ") = ", ttest_validation_den[[1]], ", p = ", ttest_validation_den[[3]],
          "\n\n",
          "Testing generalization results between density conditions against zero:\n",
          "t(", ttest_generalization_den[[2]], ") = ", ttest_generalization_den[[1]], ", p = ", ttest_generalization_den[[3]],
          sep = ""))



kable(anova_vel, caption = "Likelihood-Ratio Test: comparing information quality between Velocity variants", digits = 3)
kable(pairwise_comparison_vel$contrasts,
      caption = "Pairwise comparisons for velocity (Tukey corrected)",
      digits = 3)

kable(anova_den, caption = "Likelihood-Ratio Test: comparing information quality between Density variants", digits = 3)
kable(pairwise_comparison_den$contrasts,
      caption = "Pairwise comparisons for density (Tukey corrected)",
      digits = 3)




# Save tables as dataframes
table_anova_vel <- as.data.frame(anova_vel)
table_pairwise_vel <- as.data.frame(pairwise_comparison_vel$contrasts)
table_anova_den <- as.data.frame(anova_den)
table_pairwise_den <- as.data.frame(pairwise_comparison_den$contrasts)


saveRDS(
  list(anova_vel = table_anova_vel,
       pairwise_vel = table_pairwise_vel,
       anova_den = table_anova_den,
       pairwise_den = table_pairwise_den
  ),
  file = file.path(dir.output, "tables", "tables_prediction_velden.rds")
)


```



\newpage

## Prediction accuracies between velocity levels 

```{r plotting_predictions_vel, fig.cap = "Participant-wise averaged accuracy scores (z-standardized correlation coefficients between predicted and true PoTJs) for cross-validation (blue) and generalization (orange). For cross-validation, test-sets were composed of snapshots from the same Velocity variants. For generalization, we tested whether predictors trained on one variant of Velocity could predict PoTJs in the other Velocity variants. Left most column: averaged for all cross-validations and generalizations; Right columns: separate results for cross-validations within individual Velocity variants and averages for generalizations from ‘zero’ to ‘slow’/’fast’, from ‘slow’ to ‘zero’/’fast’ and from ‘fast’ to ‘zero’/’slow’. Each participants averages are depicted as small, transparent circles. Boxplot showing group median (middle line), 25% - 75% quantile (hinges) and the smallest value below 25% (lower whisker) and the largest value above the 75% quantile (upper whiskers) that are within 1.5 of the inter quartile range (distance between quantile 25% and 75%)."}

# ------
# Plotting accuracies of validation / generalization between velocity levels

# Creating one dataset for plotting
sep_data = cbind(data_vel, "level" = "Velocity level")
avg_data = cbind(data_vel_avg, "level" = "Average")
avg_data$vel = rep("", nrow(avg_data))
data_vel_all = rbind(sep_data, avg_data)
data_vel_all$level = factor(data_vel_all$level, levels = c("Average", "Velocity level"))


# Plot average accuracy of validation and generalization for velocity
predictionacc_vel_plot <- ggplot(data_vel_all, aes(x=vel, y=z, color = type, fill = type)) +

  geom_point(shape = 16, size = 2, alpha = .3, position = position_dodge(width = dodge_val_gen), na.rm=TRUE) +
  geom_boxplot(size = 0.5, alpha = 0, width = .2, position = position_dodge(width = dodge_val_gen), na.rm=TRUE)+
  stat_summary(
    data = subset(data_vel_all, level != "Average"),
    fun = median,
    geom = 'line',
    aes(group = type, colour = type),
    position = position_dodge(width = 0.15),
    na.rm = TRUE
  )+
  geom_hline(yintercept=0, linetype="dashed", color = "darkgrey", linewidth = 0.5) +
  ylim(-0.2,1) +
  scale_fill_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  scale_color_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  ylab("Accuracy score (z-score)") +
  facet_grid(~ level, switch = "x", scales = "free_x", space = "free_x")+
  theme_grid+
  theme(axis.title.x = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none",
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing.y = unit(0, "lines"),
        strip.text.x = element_text(margin = margin(t = 2, b = 2)),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.switch.pad.grid = unit(2, "mm")
)

ggsave(plot = predictionacc_vel_plot,
       filename = "fig6b_predictionacc_vel.png", path = dir.output,
       width = fig_width, height = fig_height, units = "mm", dpi = dpi_pngs)

predictionacc_vel_plot
```

**$\Rightarrow$ Oscillation-based prediction of PoT are independent of velocity**

\newpage

## Prediction accuracies between density levels 

```{r plotting_predictions_den, fig.cap = "Participant-wise averaged accuracy scores (z-standardized correlation coefficients between predicted and true PoTJs) for cross-validation (blue) and generalization (orange). For cross-validation, test-sets were composed of snapshots from the same Density variants. For generalization, we tested whether predictors trained on one variant of Density could predict PoTJs in the other Density variants. Left most column: averaged for all cross-validations and generalizations; Right columns: separate results for cross-validations within individual Density variants and for generalizations from few to many and from many to few. Each participants averages are depicted as small, transparent circles. Boxplot showing group median (middle line), 25% - 75% quantile (hinges) and the smallest value below 25% (lower whisker) and the largest value above the 75% quantile (upper whiskers) that are within 1.5 of the inter quartile range (distance between quantile 25% and 75%)."}

# ------
# Plotting accuracies of validation / generalization between density levels

# Creating one dataset for plotting
sep_data = cbind(data_den, "level" = "Density level")
avg_data = cbind(data_den_avg, "level" = "Average")
avg_data$den = rep("", nrow(avg_data))
data_den_all = rbind(sep_data, avg_data)
data_den_all$level = factor(data_den_all$level, levels = c("Average", "Density level"))

# Plot average accuracy of validation and generalization for density
predictionacc_den_plot <- ggplot(data_den_all, aes(x=den, y=z, color = type, fill = type)) +

  geom_point(shape = 16, size = 2, alpha = .3, position = position_dodge(width = dodge_val_gen), na.rm=TRUE) +
  geom_boxplot(size = 0.5, alpha = 0, width = .2, position = position_dodge(width = dodge_val_gen), na.rm=TRUE)+
  stat_summary(
    data = subset(data_den_all, level != "Average"),
    fun = median,
    geom = 'line',
    aes(group = type, colour = type),
    position = position_dodge(width = 0.15),
    na.rm = TRUE
  )+
  geom_hline(yintercept=0, linetype="dashed", color = "darkgrey", linewidth = 0.5) +
  ylim(-0.2,1) +
  scale_fill_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  scale_color_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  ylab("Accuracy score (z-score)") +
  facet_grid(~ level, switch = "x", scales = "free_x", space = "free_x")+
  theme_grid+
  theme(axis.title.x = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none",
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing.y = unit(0, "lines"),
        strip.text.x = element_text(margin = margin(t = 2, b = 2)),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.switch.pad.grid = unit(2, "mm"))

ggsave(plot = predictionacc_den_plot,
       filename = "fig6d_predictionacc_den.png", path = dir.output,
       width = fig_width * 0.8, height = fig_height, units = "mm", dpi = dpi_pngs)

predictionacc_den_plot
```

**$\Rightarrow$ Oscillation-based prediction of PoT are independent of density**
\newpage


## Prediction of PoT between velocity-density combinations 

```{r predictions_between_level_comb, comment = ""}

# load dataframe 'data_velden' containing validation & generalization between velocity-densiyt combinations 
load(file = file.path(dir.data, "data_velden.Rda"))

# create dataframe with averages over all velocity-density combinations
data_velden_avg = data_velden %>%
  dplyr::group_by(subject, type) %>%
  dplyr::summarize(z = mean(z, na.rm = TRUE), .groups = "drop")

data_velden_avg$velden = "total"



# =====
# Analysis

# compute average and perform t-test for validation & generalization
ttest_validation_comb = t.test(data_velden_avg[data_velden_avg$type == "validation",]$z)
ttest_generalization_comb = t.test(data_velden_avg[data_velden_avg$type == "generalization",]$z)

# analyse differential effects of velocity-density combinations
lm_velden0 = lmer(z ~ 1 + (1 | subject), subset(data_velden, type == "generalization"), REML = FALSE)
lm_velden1 = lmer(z ~ factor(vel) + (1 | subject), subset(data_velden, type == "generalization"), REML = FALSE)
lm_velden2 = lmer(z ~ factor(vel) + factor(den) + (1 | subject), subset(data_velden, type == "generalization"), REML = FALSE)
lm_velden3 = lmer(z ~ factor(vel) * factor(den) + (1 | subject), subset(data_velden, type == "generalization"), REML = FALSE)
anova_comb = anova(lm_velden0, lm_velden1, lm_velden2, lm_velden3)

# Conduct pairwise comparison between velocity-density combinations (tukey-test, 'emmeans'-package) --> suppl
pairwise_comparison_comb = lsmeans(lm_velden3, pairwise ~ vel:den)


# =====
# Printing

cat(paste("Testing cross-validation results within velocity conditions against zero:\n",
          "t(", ttest_validation_comb[[2]], ") = ", ttest_validation_comb[[1]], ", p = ", ttest_validation_comb[[3]],
          "\n\n",
          "Testing generalization results between velocity conditions against zero:\n",
          "t(", ttest_generalization_comb[[2]], ") = ", ttest_generalization_comb[[1]], ", p = ", ttest_generalization_comb[[3]],
          sep = ""))


kable(anova_comb, caption = "Likelihood-Ratio Test: comparing information quality between Velocity*Density combinations", digits = 3)
kable(pairwise_comparison_comb$contrasts,
      caption = "Pairwise comparisons for velocity-density combinations (Tukey corrected)",
      digits = 3)



# Save tables as dataframes
table_anova <- as.data.frame(anova_comb)
table_pairwise <- as.data.frame(pairwise_comparison_comb$contrasts)

saveRDS(
  list(anova = table_anova,
       pairwise = table_pairwise
  ),
  file = file.path(dir.output, "tables", "tables_prediction_comb.rds")
)


```

\newpage


```{r, plotting_predictions_level_comb, fig.cap = "Participant-wise averaged accuracy scores (z-standardized correlation coefficients between predicted and true PoTJs) for cross-validation (blue) and generalization (orange) between variants of different Velocity and Density. For cross-validation, test-sets were composed of the same starfield variant. For generalization, we tested whether predictors trained with one variant could also predict PoTJs in the other variants. Left most column: averaged for all cross-validations and generalizations; Right columns: separate results for cross-validations within variants (blue) and averages for generalizations from each specific variant to all other variants (orange). Each participants averages are depicted as small, transparent circles. Boxplot showing group median (middle line), 25% - 75% quantile (hinges) and the smallest value below 25% (lower whisker) and the largest value above the 75% quantile (upper whiskers) that are within 1.5 of the inter quartile range (distance between quantile 25% and 75%)."}


# ------
# Plotting accuracies of validation / generalization between density levels

# Creating one dataset for plotting

# Change factor labels
sep_data = cbind(data_velden, "level" = paste("Density:", as.character(data_velden$den)))
sep_data = subset(sep_data, select = -c(den))

levels(sep_data$vel) <- c("Velocity: zero", "Velocity: slow", "Velocity: fast")

avg_data = cbind(data_velden_avg, "level" = "Average")
avg_data = subset(avg_data, select = -c(velden))
avg_data$vel = rep("", nrow(avg_data))

data_velden_all = rbind(sep_data, avg_data)

data_velden_all$level = factor(data_velden_all$level, ordered = TRUE, levels = c("Average", "Density: few", "Density: many"))


# Plot average accuracy of validation and generalization for velocity
predictionacc_velden_plot <- ggplot(data_velden_all, aes(x=vel, y=z, color = type, fill = type)) +

  geom_point(shape = 16, size = 2, alpha = .3, position = position_dodge(width = 0.35), na.rm=TRUE) +
  geom_boxplot(size = 0.5, alpha = 0, width = .3, position = position_dodge(width = 0.35), na.rm=TRUE)+
  stat_summary(
    data = subset(data_velden_all, level != "Average"),
    fun = median,
    geom = 'line',
    aes(group = type, colour = type),
    position = position_dodge(width = 0.35),
    na.rm = TRUE
  )+
  geom_hline(yintercept=0, linetype="dashed", color = "darkgrey", linewidth = 0.5) +
  ylim(-0.2,1) +
  scale_fill_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  scale_color_manual(values = c("validation" = colors.val_gen[1],
                                "generalization" = colors.val_gen[2]))+
  ylab("Accuracy score (z-score)") +
  facet_grid(~ level, switch = "x", scales = "free_x", space = "free_x")+
  theme_grid+
  theme(axis.text.x = element_text(size = 7),
        axis.title.x = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none",
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing.y = unit(0, "lines"),
        strip.text.x = element_text(margin = margin(t = 2, b = 2)),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.switch.pad.grid = unit(2, "mm"))


ggsave(plot = predictionacc_velden_plot,,
       filename = "fig7b_predictionacc_velden.png", path = dir.output,
       width = fig_width, height = fig_height, units = "mm", dpi = dpi_pngs)

predictionacc_velden_plot
```

**$\Rightarrow$ Oscillation-based prediction of PoT are independent of moment-to-moment change rate**

